### 2017-03-03 paper/chapters/vae.tex
last-minute changes

Using gradient-based optimization techniques enables the encoder network to be trained to output distributions similar to the true posterior $p(z|x)$.
In addition to this inference network there is a generator network which transforms points $z$ in the latent space back to the data space while trying to minimize the reconstruction error.

@@

Such function approximator will emit the variational parameters $\phi$ of the probability encoder distribution $q_\phi(z|x)$.
Equally for the probabilistic decoder which constructs a distribution $p_\theta(x|z)$ with the parameters $\theta$.
This architecture is shown in Figure \ref{fig:vae_architecture} with a one-dimensonal gaussian prior.
\\

@@

\subsection{Approximated Inference}
A problem which often arises in directed generative models is the \emph{inference problem}.
As discussed in the introduction~\ref{sec:introduction}, we assume that the observations are stochastically dependent on some latent variables.
Inference is called the process of coming up with values of $z$ which can produce the observations, which formally means evaluating $p(z|x)$ where $z$ are the latent variables and $x$ observations.
\begin{equation}
  p(z|x) = \frac{p(x|z)p(z)}{p(x)}\\
  \label{eq:intractable_posterior}
\end{equation}
where
\begin{conditions}
  p(z|x) & posterior distribution\\
  p(x|z) & likelihood of $x$ given $z$\\
  p(z)   & prior distribution of $z$\\
\end{conditions}
Using Bayes's rule we can expand the posterior distribution $p(z|x)$ as shown in Equation~\ref{eq:intractable_posterior}.
Note that to evaluate the denominator $p(x)$, all hidden variables need to be marginalized: $p(x) = \int p(x,z) dz$.\\
Due to the possibly large number of hidden variables, this integral can be high-dimensional and therefore difficult to calculate.
This integral can be difficult to compute and needs to be approximated.
However, for a large number of hidden variables, this integral is intractable to compute exactly and therefore needs to be approximated~\cite[p.~461-462]{bishop:2006}.
In order to approximate those intractable posterior distributions, we discuss two commonly used techniques:  Markov chain Monte Carlo and variational inference.

@@

\paragraph{Variational Inference}
% variational inference
Due to the possibly slow convergence of MCMC methods, variational inference (VI)~\cite[Chapter~19.4]{deeplearning:2016} is used in the variational autoencoder.
VI works by picking a specific distribution restricted to a family of distributions which matches the posterior most closely. This means that VI is used to turn the inference problem into an optimization technique, allowing for gradient-based optimization.
%TODO: more on advantages and disadvantages
Using VI yields better performance at the sacrifice of theoretical guarantee of matching the posterior distribution.

More specifically, VI turns the inference problem of computing $p(z|x)$ where $z$ are the hidden and $x$ the observed variables into an optimization problem, where we can use gradient-based methods.

\subsubsection{Kullback-Leibler Divergence}
In order to approximate one distribution $Q$ to another distribution $P$, there needs to be a measurement of how close $Q$ is to $P$.
One commonly used measurement is called Kullback-Leibler (KL) divergence\cite{kl_div:1951} which enables the variational inference methods to choose appropiate distributions.
%\paragraph{The Kullback-Leibler Divergence} (KL divergence) is a measurement for the difference between two probability distributions. $\mathcal{D}_{\mathrm{KL}}(P \| Q)$ can informally described as the amount of information which is lost when using $Q$ to represent $P$.
The KL divergence from $Q$ to $P$ is defined in Equation \ref{eq:kl_div}.
\begin{equation}
  \label{eq:kl_div}
  \mathcal{D}_{\mathrm{KL}}(P || Q) = \mathbb{E}_{x \sim P(x)}\bigg[log \frac{P(x)}{Q(x)}\bigg]
\end{equation}
\todo[inline]{Add some description to the equation above, logarithmic difference between P and Q, bla bla expectation}
%For discrete probability distributions $P$ and $Q$, the KL divergence from $Q$ to $P$ is defined\cite{kl_div:1951} in equation \ref{eq:kl_div_disc}:
%\begin{equation}
  %\label{eq:kl_div_disc}
  %\mathcal{D}_{\mathrm{KL}}(P || Q) = \sum_{i} P(i) \log \frac{P(i)}{Q(i)}
%\end{equation}
%For continuous distributions $P$ and $Q$ the KL divergence is defined similarly in equation \ref{eq:kl_div_cont}.
%\begin{equation}
  %\label{eq:kl_div_cont}
  %\mathcal{D}_{\mathrm{KL}}(P || Q) = \int P(i) \log \frac{P(i)}{Q(i)} di
%\end{equation}

@@

However we are not able to optimize Equation \ref{eq:vae_obj} directly, as the intractable evidence $p(x)$ appears during the expansion of the posterior $p(z|x)$ to $\frac{p(x|z)p(z)}{p(x)}$.\\
The denominator $p(x)$ can be computed by marginalizing out the hidden variables $z$: 
\begin{equation}
  \label{eq:integral}
p(x) = \int p(x,z) p(z) dz
\end{equation}

@@

\paragraph{Regularization term $\mathcal{L}_{reg}$} encourages the model to learn simple representations in latent space, due to the negative KL divergence between the learned variational distribution $q_\phi(z|x)$ and the latent space prior distribution $p(z)$.

@@

Using the example of handwritten digits it appears why this regularizer term $\mathcal{L}_{reg}$ is important for the model to learn generalized representations in latent space.
Suppose we have two data points from the dataset which both represent the same handwritten digit.
Ideally, our model would learn representations in latent space to be fairly close to ensure the model has learned to generalize the semantics. Without $\mathcal{L}_{reg}$ the model would assign each data example a different representation in latent space with a distribution collapsing to that single point. This behaviour is undesired and is as such penalitized.

%Suppose two data points with similar input
%Without $\mathcal{L}_{reg}$ the encoder would encode data points we want to have similar representations, fo

@@

\vspace{2cm}

\paragraph{Reconstruction error $\mathcal{L}_{rec}$} measures how well the decoder reconstructs the input data using the latent space $z$.
This error is measured using the negative log-likelihood of the conditional probability distribution $p_\theta(x|z)$ where $z$ is sampled from $q_\phi(z | x)$.\\
$\mathcal{L}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.

@@

% more math:
% http://stats.stackexchange.com/questions/199605/how-does-the-reparameterization-trick-for-vaes-work-and-why-is-it-important
To be able to apply the backpropagation\cite{backprop:1988} algorithm to the loss function of the VAE without any injected noise it has to be both differentiable and deterministic.
However, as $z$ is sampled from the constructed probability distribution $q_\phi(z|x)$ it is stochastic.
This poses a problem, because the gradient can not be computed accurately due to the added noise.\\
In order to circumvent this restriction, the reparameterization trick (as proposed by \cite{vae:2014} and \cite{dlgm:2014}) can be applied.
Instead of drawing $z$ directly from $q_\phi(z|x)$, we first sample an auxiliary variable $\epsilon$ from the isotropic gaussian $\mathcal{N}(0, I)$. Then we apply Equation \ref{eq:rep_trick} to transform $\epsilon$ into $z$.
\begin{equation}
  \label{eq:rep_trick}
  z = \mu(x) + \Sigma^{1/2}(x)*\epsilon
\end{equation}
Note that this transformation is entirely deterministic and as such allows for backpropagation through the transformation into the variational parameters $\phi$.

Figure \ref{fig:rep_trick} expresses the reparameterization trick applied to an arbitrary function $f$ with the parameters $\theta$.\\
Using this trick, the whole VAE model can be trained using gradient-based optimization techniques.\\\\

Note further that in most common architectures, the distribution class enforced on the latent space is gaussian, there exists similar reparameterization techniques allowing different prior probability distributions on $z$.
%The core idea of VAEs is to match a distribution $q(x)$ to the desired $p(x)$ by enforcing a lower variational bound on $q(x)$ in a way that it seeks to match $p(x)$.
%In contrast to other autoencoders, VAEs behave differently due to this lower bound but they resemble the architecture of traditional autoencoders (sparse, denoising).

@@

\subsection{Learning}
\label{sub:vae_learning}
The probabilistic encoder $q_\phi(z|x)$ and decoder $p_\theta(x|z)$ are commonly implemented using deep neural networks.
To allow analytical computation of the ELBO, the prior on the latent space $z$ is often choosen to be the isotropic gaussian $p(z) = \mathcal{N}(0,1)$.

%To allow the computation of the KL divergence between $q_\theta(z|x)$ and the latent prior distribution $p(z)$ in closed form, the prior is an isotropic gaussian $\mathcal{N}(0, 1)$.

\begin{algorithm}
  \caption{Learning in the VAE model}
  \label{alg:vae}
  \begin{algorithmic}[1]
    %\Let{$\eta_d, \eta_g$}{Learning rate of discriminator and generator networks, respectively}
    %\Let{$\theta_d, \theta_g$}{Parameters of function approximators}
    \While{has not reached convergence}
      \Let{$x$}{sample from $p_{data}(x)$}
      \Let{$\epsilon$}{Random samples from noise prior $p(\epsilon)$}
      \Let{$g$}{Compute gradients $\nabla_{\theta,\phi}\mathcal{L}(\theta, \phi; x, \epsilon)$}
      \Let{$\theta, \phi$}{Update according to gradients $g$}
    \EndWhile
  \end{algorithmic}
\end{algorithm}

\subsection{Experiments}
In order to evaluate the inference performance and explore the learned latent space, the VAE has been implemented and trained\footnote{The source code is available on GitHub: \url{https://github.com/Spotlight0xff/deeplearning-seminar}}. Results of this experiments are shown in Figure \ref{fig:vae_experiments}.
%The model has been trained on the MNIST dataset \cite{mnist} with satisfactory generative performance as shown in figure \ref{fig:vae_experiments}

@@

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[height=7cm]{media/vae_manifold.pdf}
        \caption{Learned manifold using 2-dimensional latent space}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[height=7cm]{media/vae_samples.pdf}
        \caption{Generated images by sampling from latent space}
    \end{subfigure}
    \caption{Manifold Learning in the VAE}
    \label{fig:vae_experiments}
    \medskip
    \small
    (a) Evaluation of a variational autoencoder with 2 hidden layers with each 500 nodes and a two-dimensional latent space.
    The shown manifold has been plotted using samples uniformly samples in the range $[-3,-3]^2$ and then reconstructed using the decoder network $p_\theta(x|z)$. The results have shown that the model has learned to arrange the latent space in a sensible way by placing digits with high similarity near each other.\\
    (b) generative performance qualitively evaluated using samples in the data space by random sampling $z \sim p(z)$ and decoding $z$ using the decoder network. For this experiment an 50-dimensional latent space has been used.
\end{figure*}

>>>>

Using gradient-based optimization techniques enables the encoder network to be trained to output distributions similar to the true posterior $p(z|x)$.
In addition to this inference network, there is a generator network which transforms points $z$ in the latent space back to the data space while trying to minimize the reconstruction error.

@@

Such function approximator will emit the variational parameters $\phi$ of the probability encoder distribution $q_\phi(z|x)$.
Equally for the probabilistic decoder which constructs a distribution $p_\theta(x|z)$ with the parameters $\theta$.
This architecture is shown in Figure \ref{fig:vae_architecture} with a one-dimensional Gaussian prior.
\\

@@

\subsection{Approximated Inference}
A problem which often arises in directed generative models is the \emph{inference problem}~\cite[Chapter~16.6]{deeplearning:2016}.
As discussed in \ref{sec:introduction}, we assume that the observations are stochastically dependent on some latent variables.
Inference is called the process of coming up with values of $z$ which can produce the observations, which formally means evaluating $p(z|x)$ where $z$ are the latent variables and $x$ observations.
\begin{equation}
  p(z|x) = \frac{p(x|z)p(z)}{p(x)}\\
  \label{eq:intractable_posterior}
\end{equation}
where
\begin{conditions}
  p(z|x) & posterior distribution\\
  p(x|z) & likelihood of $x$ given $z$\\
  p(z)   & prior distribution of $z$\\
\end{conditions}
Using Bayes's rule we can expand the posterior distribution $p(z|x)$ as shown in Equation~\ref{eq:intractable_posterior}.
Note that to evaluate the denominator $p(x)$, all hidden variables need to be marginalized: $p(x) = \int p(x,z) dz$.\\
Due to the possibly large number of hidden variables, this integral can be high-dimensional and therefore difficult to calculate.
This integral can be difficult to compute and needs to be approximated.
However, for a large number of hidden variables, this integral is intractable to compute exactly and therefore needs to be approximated~\cite[p.~461-462]{bishop:2006}.
In order to approximate those intractable posterior distributions, we discuss two commonly used techniques:  Markov chain Monte Carlo and variational inference.

@@

\paragraph{Variational Inference}
% variational inference
Due to the possibly slow convergence of MCMC methods, variational inference (VI)~\cite[Chapter~19.4]{deeplearning:2016}\cite[Chapter~10.1]{bishop:2006} is used in the variational autoencoder~\cite{vae:2014}.
VI works by picking a specific distribution restricted to a family of distributions which matches the posterior most closely.
In the case of the VAE, we use neural networks to select the parameters of the distributions to turn the inference problem into an optimization problem, allowing for gradient-based optimization.
Due to the restriction of VI to a specific probability distribution class and neural networks as non-convex function approximators, it is not guaranteed to converge.
%Variational inference is however restricted to the specific class of distributions, which means that while we are able to approximate distribution in a given family (e.g. normal distributions), if the target distribution is not in that family there is no way of convergence.
%TODO: more on advantages and disadvantages
Conclusively, even though MCMC offers approximation with guaranteed convergence~\cite[Chapter~11.2]{bishop:2006}, in practice it is often slow to converge while VI offers reasonable good approximation at greater speed~\cite{vae:2014}.

\subsubsection{Kullback-Leibler Divergence}
In order to approximate one distribution $Q$ to another distribution $P$, we need to have a measurement of how close $Q$ is to $P$.
One commonly used measurement is called Kullback-Leibler (KL) divergence~\cite{kl_div:1951} which enables the variational inference methods to choose appropriate distributions.
%\paragraph{The Kullback-Leibler Divergence} (KL divergence) is a measurement for the difference between two probability distributions. $\mathcal{D}_{\mathrm{KL}}(P \| Q)$ can informally described as the amount of information which is lost when using $Q$ to represent $P$.
The KL divergence from $Q$ to $P$ is defined in Equation \ref{eq:kl_div}.
\begin{equation}
  \label{eq:kl_div}
  \mathcal{D}_{\mathrm{KL}}(P || Q) = \mathbb{E}_{x \sim P(x)}\bigg[log \frac{P(x)}{Q(x)}\bigg]
\end{equation}
%\todo{Add some description to the equation above, logarithmic difference between P and Q, bla bla expectation}
%For discrete probability distributions $P$ and $Q$, the KL divergence from $Q$ to $P$ is defined\cite{kl_div:1951} in equation \ref{eq:kl_div_disc}:
%\begin{equation}
  %\label{eq:kl_div_disc}
  %\mathcal{D}_{\mathrm{KL}}(P || Q) = \sum_{i} P(i) \log \frac{P(i)}{Q(i)}
%\end{equation}
%For continuous distributions $P$ and $Q$ the KL divergence is defined similarly in equation \ref{eq:kl_div_cont}.
%\begin{equation}
  %\label{eq:kl_div_cont}
  %\mathcal{D}_{\mathrm{KL}}(P || Q) = \int P(i) \log \frac{P(i)}{Q(i)} di
%\end{equation}

@@

However, we are not able to optimize Equation \ref{eq:vae_obj} directly, as the intractable evidence $p(x)$ appears during the expansion of the posterior $p(z|x)$ to $\frac{p(x|z)p(z)}{p(x)}$.\\
The denominator $p(x)$ can be computed by marginalizing out the hidden variables $z$: 
\begin{equation}
  \label{eq:integral}
p(x) = \int p(x,z) p(z) dz
\end{equation}

@@

% excluded due to no references found :/
%\paragraph{Regularization term $\mathcal{L}_{reg}$} encourages the model to learn simple representations in latent space, due to the negative KL divergence between the learned variational distribution $q_\phi(z|x)$ and the latent space prior distribution $p(z)$.

@@

%Using the example of handwritten digits it appears why this regularizer term $\mathcal{L}_{reg}$ is important for the model to learn generalized representations in latent space.
%Suppose we have two data points from the dataset which both represent the same handwritten digit.
%Ideally, our model would learn representations in latent space to be fairly close to ensure the model has learned to generalize the semantics. Without $\mathcal{L}_{reg}$ the model would assign each data example a different representation in latent space with a distribution collapsing to that single point. This behaviour is undesired and is as such penalitized.

%%Suppose two data points with similar input
%Without $\mathcal{L}_{reg}$ the encoder would encode data points we want to have similar representations, fo

@@

%\vspace{2cm}

%\paragraph{Reconstruction error $\mathcal{L}_{rec}$} measures how well the decoder reconstructs the input data using the latent space $z$.
%This error is measured using the negative log-likelihood of the conditional probability distribution $p_\theta(x|z)$ where $z$ is sampled from $q_\phi(z | x)$.\\
%$\mathcal{L}_{\mathrm{rec}}$ is necessary to force the encoder to produce latent variables which can be used to reconstruct the input data well.

@@

% more math:
% http://stats.stackexchange.com/questions/199605/how-does-the-reparameterization-trick-for-vaes-work-and-why-is-it-important
To be able to apply the backpropagation\cite{backprop:1988} algorithm to the loss function of the VAE without any injected noise it has to be both differentiable and deterministic.
However, as $z$ is sampled from the constructed probability distribution $q_\phi(z|x)$ it is stochastic.
This poses a problem because the gradient can not be computed accurately due to the added noise.\\
In order to circumvent this restriction, the reparameterization trick~\cite{vae:2014}\cite{dlgm:2014}\cite[chapter~20.9]{deeplearning:2016} can be applied.
Instead of drawing $z$ directly from $q_\phi(z|x)$, we first sample an auxiliary variable $\epsilon$ from the isotropic gaussian $\mathcal{N}(0, I)$. Then we apply Equation \ref{eq:rep_trick} to transform $\epsilon$ into $z$.
\begin{equation}
  \label{eq:rep_trick}
  z = \mu(x) + \Sigma^{1/2}(x)*\epsilon
\end{equation}
Note that this transformation is entirely deterministic and as such allows for backpropagation through the transformation into the variational parameters $\phi$.

Figure \ref{fig:rep_trick} expresses the reparameterization trick applied to an arbitrary function $f$ with the parameters $\theta$.\\
Using this trick, the whole VAE model can be trained using gradient-based optimization techniques.

%Note further that in most common architectures, the distribution class enforced on the latent space is gaussian, there exists similar reparameterization techniques allowing different prior probability distributions on $z$.

%The core idea of VAEs is to match a distribution $q(x)$ to the desired $p(x)$ by enforcing a lower variational bound on $q(x)$ in a way that it seeks to match $p(x)$.
%In contrast to other autoencoders, VAEs behave differently due to this lower bound but they resemble the architecture of traditional autoencoders (sparse, denoising).

@@

\subsection{Learning}
\label{sub:vae_learning}
The probabilistic encoder $q_\phi(z|x)$ and decoder $p_\theta(x|z)$ are commonly implemented using deep neural networks.
To allow analytical computation of the ELBO, the prior on the latent space $z$ is often choosen to be the isotropic gaussian $p(z) = \mathcal{N}(0,1)$.
The learning procedure of a variational autoencoder is shown in Algorithm~\ref{alg:vae} which is a simplified version of the algorithm in the original paper~\cite{vae:2014}.
\begin{algorithm}
  \caption{Learning in the VAE model}
  \label{alg:vae}
  \begin{algorithmic}[1]
    %\Let{$\eta_d, \eta_g$}{Learning rate of discriminator and generator networks, respectively}
    %\Let{$\theta_d, \theta_g$}{Parameters of function approximators}
    \While{has not reached convergence}
      \Let{$x$}{sample from $p_{data}(x)$}
      \Let{$\epsilon$}{Random samples from noise prior $p(\epsilon)$}
      \Let{$g$}{Compute gradients $\nabla_{\theta,\phi}\mathcal{L}(\theta, \phi; x, \epsilon)$}
      \Let{$\theta, \phi$}{Update according to gradients $g$}
    \EndWhile
  \end{algorithmic}
\end{algorithm}
%
\subsection{Experiments}
In order to evaluate the inference performance and explore the learned latent space, the VAE has been implemented and trained using TensorFlow~\cite{tf:2015} on the MNIST dataset implementing a fully-connected network.%\footnote{The source code is available on GitHub: \url{https://github.com/Spotlight0xff/deeplearning-seminar}}.
Both experiments as shown in Figure~\ref{fig:vae_experiments} have been performed with two hidden layers of size 500.
Rectified linear units~\cite{relu:2000} were used as activation functions in this network.

%The model has been trained on the MNIST dataset \cite{mnist} with satisfactory generative performance as shown in figure \ref{fig:vae_experiments}

@@

\begin{figure*}[h!]
    \centering
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[height=7cm]{media/vae_manifold.pdf}
        \caption{Learned manifold}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[height=7cm]{media/vae_samples.pdf}
        \caption{Randomly generated images by sampling}
    \end{subfigure}
    \caption{Manifold Learning in the VAE}
    \label{fig:vae_experiments}
    \medskip
    \small
    (a) Evaluation of a variational autoencoder with 2 hidden layers with each 500 nodes and a two-dimensional latent space.
    The shown manifold has been plotted using samples uniformly samples in the range $[-3,-3]^2$ and then reconstructed using the decoder network $p_\theta(x|z)$.
    (b) generative performance qualitatively evaluated using samples in the data space by random sampling $z \sim p(z)$ and decoding $z$ using the decoder network. For this experiment a 50-dimensional latent space has been used.
\end{figure*}

### 2017-03-03 paper/chapters/vae/cvae.tex
last-minute changes

Instead of using VAEs in a completely unsupervised manner on a unlabeled dataset, it is also possible to learn from datasets where a small subset of datapoints has corresponding labels while most datapoints are still unlabeled.
Techniques learning from such datasets are called semi-supervised learning and can improve performance significantly~\cite{cvae:2014}\cite{cvae_2:2015}.\\
%
To adjust the VAE model to the semi-supervised case, we have to add a random variable $y$ corresponding to the label if available.\\\\
First, instead of using samples from the dataset alone we are now also conditioning on $y$, formally 
$x \sim p_\theta(x|y,z)$ instead of $x \sim p_\theta(x|z)$.\\
Then the probabilistic encoder and decoder can be written as $q_\theta(z|x,y)$ and $p_\phi(x|y,z)$ respectively.\\
%
%The modified setup of the semi-supervised case for the variational autoencoder can be formalized as follows.
%Dataset $X$ of size $N$ with observations $x_i$ with $i<N$ where for some small subset of observations $x_i$ the corresponding labels $y_i$ are known.
%We denote the labelled subset as $~p_l(x,y)$ and the unlabelled dataset as $~p_u(x,y)$.
%\begin{equation}
%
%\end{equation}
%Instead of randomly generated data, generation of output conditioned on some prior information is needed.
%A few examples would be super-resolution of images (citation!), prediction of subsequent frames of an video or other prediction tasks
%
%Conditional VAE is a modification of the original VAE framework to support conditioning on another variable $y$ in addition to the input data $x$.
%
%To support conditional distributions, the following relations have to be rewritten:
%
These modifications allows us to reformulate the objective function $\mathcal{L}$ from Equation~\ref{eq:vae_obj} in equations \ref{eq:cvae_labels} and \ref{eq:cvae_unknown}.

@@

The final objective function for whole dataset is the summation between right-hand terms of equations \ref{eq:cvae_labels} for the labelled subset and \ref{eq:cvae_unknown} for the unlabelled subset.

>>>>

Instead of using VAEs in a completely unsupervised manner on an unlabeled dataset, it is also possible to learn from datasets where a small subset of data points has corresponding labels while most data points are still unlabeled.
Techniques learning from such datasets are called semi-supervised learning and can improve performance significantly~\cite{cvae:2014}\cite{cvae_2:2015}.\\
%
To adjust the VAE model to the semi-supervised case, we have to add a random variable $y$ corresponding to the label if available.\\\\
First, instead of using samples from the dataset alone we are now also conditioning on $y$, formally 
$x \sim p_\theta(x|y,z)$ instead of $x \sim p_\theta(x|z)$.\\
Then the probabilistic encoder and decoder can be written as $q_\theta(z|x,y)$ and $p_\phi(x|y,z)$ respectively.\\
%
%The modified setup of the semi-supervised case for the variational autoencoder can be formalized as follows.
%Dataset $X$ of size $N$ with observations $x_i$ with $i<N$ where for some small subset of observations $x_i$ the corresponding labels $y_i$ are known.
%We denote the labelled subset as $~p_l(x,y)$ and the unlabelled dataset as $~p_u(x,y)$.
%\begin{equation}
%
%\end{equation}
%Instead of randomly generated data, generation of output conditioned on some prior information is needed.
%A few examples would be super-resolution of images (citation!), prediction of subsequent frames of an video or other prediction tasks
%
%Conditional VAE is a modification of the original VAE framework to support conditioning on another variable $y$ in addition to the input data $x$.
%
%To support conditional distributions, the following relations have to be rewritten:
%
These modifications allows us to reformulate the objective function $\mathcal{L}$ from Equation~\ref{eq:vae_obj} in equations \ref{eq:cvae_labels} and \ref{eq:cvae_unknown}.

@@

The final objective function forthe  whole dataset is the summation between right-hand terms of equations \ref{eq:cvae_labels} for the labeled subset and \ref{eq:cvae_unknown} for the unlabelled subset.

### 2017-03-03 paper/chapters/vae/iwae.tex
remove dumb emphasis
last-minute changes

% write about IWAE, extends these k passes with importance weights
Importance Weighted Autoencoders (IWAE\cite{iwae:2015}) extends on the VAE framework using $k$-sampled importance weighting of the log-likelihood.
% add intuition for IWAE, ruslan explained it very well during his DL summer school talk! (talk_Montreal_2016_Salakhutdinov.pdf)
Using importance weighting can be informally interpreted as putting more weight on data with high likelihood.

@@

Additionally, if $\frac{p(h,x)}{q(h|x)}$ is bounded, then the model is guaranteed to reach the optimimum as $k$ approaches infinity in equation \ref{eq:iwae}
\begin{equation}
\label{eq:iwae}
\lim_{k \rightarrow \infty} \mathcal{L}_{k}(x) = \log p(x)
\end{equation}

>>>>

% write about IWAE, extends these k passes with importance weights
The Importance Weighted Autoencoder\cite{iwae:2015} (IWAE) extends on the VAE framework using $k$-sampled importance weighting of the log-likelihood.
% add intuition for IWAE, ruslan explained it very well during his DL summer school talk! (talk_Montreal_2016_Salakhutdinov.pdf)
Using importance weighting can be informally interpreted as putting more weight on data with high likelihood.

@@

Additionally, if $\frac{p(h,x)}{q(h|x)}$ is bounded, then the model is guaranteed to reach the optimum as $k$ approaches infinity in equation \ref{eq:iwae}
\begin{equation}
\label{eq:iwae}
\lim_{k \rightarrow \infty} \mathcal{L}_{k}(x) = \log p(x)
\end{equation}
